---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

R package: `md.tools`
=====================

<!-- badges: start -->
<!-- badges: end -->

A miscellaneous set of tools for working with *masked data* and common
features of masked data.
The tool set takes inspiration from functional programming, with inputs and
outputs defined over masked data frames of type `tbl_md` (or just data frames),
making it consistent with the *tidyverse* way of doing things.

We provide a set of simple functions on masked data frames, which
may be used to compose more complicated functions, particularly when using
the pipe operator `%>%`.

#### Installation

You can install the development version of `md.tools` from
[GitHub](https://github.com/queelius/md.tools) with:

```{r,message=F,warning=F}
# install.packages("devtools")
devtools::install_github("queelius/md.tools")
library(tidyverse)
library(md.tools)
```

## Matrices

A lot of space in `md.tools` is devoted to working with matrices encoded in
the columns of data frames.
We could directly store matrices in a column, but we prefer to work with columns
defined over primitive types like `boolean`.

Consider the `boolean` matrix `C` of size `10`-by-`3`:
```{r}
C <- matrix(sample(c(T,F), size=3*10, replace=TRUE), nrow=10)
```

We may represent this in a data frame of `10` rows with the columns `c.1`,
`c.2`, and `c.3` with:
```{r}
md <- md_encode_matrix(C,"c")
print(md)
```

We may also decode a matrix stored in a data frame with:
```{r}
C.decoded <- md_decode_matrix(md,"c")
print(all(C == C.decoded))
```

```{r}
m <- ncol(C)
n <- nrow(C)

md$k <- sample(1:m,n,replace=TRUE)
#md <- md_cand_contains(md)
print(md)
```

## Metadata
We prefer to work with plaintext data files.
To store data frames, we prefer to work with plaintext files like CSV files,
where each row corresponds to some set of measurements of some experimental
unit.

However, we may also want to store *metadata* about the experiment that
generated the data, or we may wish to store more information about the
experimental units that does not naturally fit into the data frame model.

To store metadata, we take the general approach of storing JSON (Javscript
Object Notation) in the *comments* of the tabular data file (like CSV), where
a comment by default is anything after the `#` character on a line.

```{r}
md <- md.tools::md_read_csv_with_meta("./raw-data/exp_series_data_1.csv")
print(md)
```
